---
title: "Project4-A2+KNN"
output:
  html_document:
    df_print: paged
  pdf_document: default
editor_options: 
  chunk_output_type: console
---


### Step 1 Load Data and Train-test Split
```{r}
library(dplyr)
library(tidyr)
library(MASS)
library(ggplot2)
data <- read.csv("../data/ml-latest-small/ratings.csv")
set.seed(0)
test_idx <- sample(1:nrow(data), round(nrow(data)/5, 0))
train_idx <- setdiff(1:nrow(data), test_idx)
data_train <- data[train_idx,]
data_test <- data[test_idx,]
```

###Step 2 Matrix Factorization
#### Step 2.1 Algorithm and Regularization

A2. [Gradient Descent with Probabilistic Assumptions](./paper/P3 probabilistic-matrix-factorization.pdf) Section 2


```{r}
U <- length(unique(data$userId))
I <- length(unique(data$movieId))
source("../lib/Matrix_Factorization_pmf.R")
#RMSE function
RMSE <- function(rating, est_rating){
  sqr_err <- function(obs){
    sqr_error <- (obs[3] - est_rating[as.character(obs[2]), as.character(obs[1])])^2
    return(sqr_error)
  }
  return(sqrt(mean(apply(rating, 1, sqr_err))))  
}
```




#### Step 2.2 Parameter Tuning

```{r}
source("../lib/cross_validation_pmf.R")
f_l = cbind(c(10,20,10,20),c(1,1,0.5,0.5),c(1,1,0.5,0.5))
```

```{r, eval=FALSE}
result_summary <- array(NA, dim = c(nrow(f_l), 5, 4)) 
run_time <- system.time(for(i in 1:nrow(f_l)){
    par <- paste("f = ", f_l[i,1], ", sigma_p = ", f_l[i,2],", sigma_q = ", f_l[i,3])
    cat(par, "\n")
    current_result <- cv.function_pmf(data, K = 5, f = f_l[i,1],sigma_p=f_l[i,2],sigma_q= f_l[i,3],sigma = 0.1)
    result_summary[,,i] <- matrix(unlist(current_result), ncol = 5, byrow = T) 
    print(result_summary)
})
save(result_summary, file = "../output/pmf/rmse_pmf.Rdata")
```

```{r}
load("../output/pmf/rmse_pmf.Rdata")
rmse_pmf <- data.frame(rbind(t(result_summary[1,,]), t(result_summary[2,,])), train_test = rep(c("Train", "Test"), each = 4), par = rep(paste("f = ", f_l[,1], ", sigma_p = ", f_l[,2], ", sigma_q = ", f_l[,3]), times = 2)) %>% gather("epoch", "RMSE", -train_test, -par)
rmse_pmf$epoch <- as.numeric(gsub("X", "", rmse_pmf$epoch))
rmse_pmf %>% ggplot(aes(x = epoch, y = RMSE, col = train_test)) + geom_point() + facet_grid(~par)
```

### Step 3 Postprocessing
After matrix factorization, postporcessing will be performed to improve accuracy.
The referenced papers are:

P2:[Postprocessing SVD with KNN](./paper/P2 Improving regularized singular value decomposition for collaborative filtering .pdf) Section 3.5


```{r, eval= FALSE}
result <- gradesc_pmf(f = 10, sigma_p = 0.5,sigma_q = 0.5, lrate = 0.01, max.iter = 100, stopping.deriv = 0.01, data = data, train = data_train, test = data_test)

save(result, file = "../output/pmf/mat_fac_pmf.RData")
```

```{r}
load(file = "../output/pmf/mat_fac_pmf.RData")

#result[["q"]][,i] #ith movie features值
library(lsa)

cossim_mat <- cosine(as.matrix(result[["q"]]))
save(cossim_mat, file = "../output/pmf/cossim.Rdata")
closest <- rep(NA, ncol(result[["q"]]))
for (i in 1:ncol(result[["q"]])){
  x<- cossim_mat[i,]
  x<-t(x)
  closest[i] <- which(grepl(max(x[-which.max(x)]),x))# return the second highest cosine sim(drop itself)
}
```

```{r}
# calculate the mean rate for different movies
meanrate <- data%>%group_by(movieId)%>%summarize(mean_rating=mean(rating))
# median
medianrate <- data%>%group_by(movieId)%>%summarize(median_rating=median(rating))
```

```{r}
library(tidyverse)
DAT<-data
DAT <- DAT [,-4]
DAT <- DAT %>% spread(movieId, rating)
DAT1 <- data.frame(index=seq(1:5931640))
DAT1$y <-as.vector(as.matrix(DAT[,-1]))
DAT1$movieId<-rep(1:9724, times = 610)
DAT1$userId <-rep(1:610, times = 9724)


#add X2
knn_rating <- c()
for (m in 1:9724) {
  knn_rating[m] <- meanrate[closest[m], 2]
}
knn_rating<-unlist(knn_rating)
DAT1$KNN<- rep(knn_rating, times =610)

# prepare x1 predictor
q <- as.matrix(result[["q"]])
p <- as.matrix(result[["p"]])
r<-t(p)%*%q
x1<-as.vector(r)

DAT1$X1<-x1

# remove NA

DAT2<-na.omit(DAT1)
DAT2<-DAT2[,-1]

# regression model

reg<-lm(y~KNN+X1,data=DAT2)
summary(reg)
```

### Step 4 Evaluation

```{r}
library(ggplot2)

RMSE <- data.frame(epochs = seq(20, 100, 20), Training_MSE = result$train_RMSE, Test_MSE = result$test_RMSE) %>% gather(key = train_or_test, value = RMSE, -epochs)

RMSE %>% ggplot(aes(x = epochs, y = RMSE,col = train_or_test)) + geom_point() + scale_x_discrete(limits = seq(20, 100, 20)) + xlim(c(0, 100))

```
